\documentclass[12pt,reqno]{amsart}
\usepackage{./header, amssymb}

\hdr{Mathematical Statistics}{Chapter 6: Random vectors}

\begin{document}

\bigskip

\prob Suppose that we flip a loaded coin twice, with probability $p$ of landing heads. Let $X=1$ if the first flip lands heads, and $X=0$ if it lands tails. Similarly, let $Y=1$ if the second flip lands heads, and $Y=0$ if it lands tails. Compute the joint distribution of $(X,Y)$. Verify that your answer is correct by checking that all probabilities sum to $1$.\vfill














\prob Let $(X,Y)$ be discrete with probability mass function $p(x,y)$ given in the following table:


\[\begin{tabular}{c|c|c|c|c|c}
$x \diagdown y$ & 0 & 1 & 2 & 3 & 4  \\ \hline
0 & 0.08 & 0.07 & 0.06 & 0.01 & 0.01 \\ \hline
1 & 0.06 & 0.10 & 0.12 & 0.05 & 0.02 \\ \hline
2 & 0.05 & 0.06 & 0.09 & 0.04 & 0.03 \\ \hline
3 & 0.02 & 0.03 & 0.03 & 0.03 & 0.04 
\end{tabular}\]

\medskip
\begin{enumerate}
\item Compute $P(X\leq 2, Y\leq 2)$.\vfill
\item Compute $P(X=Y)$.\vfill
\item Compute $P(X>Y)$.\vfill
\end{enumerate}
















\newpage
\prob Suppose that $(X,Y)$ is continuous with probability density function

	\[
	f(x,y) = \begin{cases}
	cx^2 y & : x^2 \leq y \leq 1, \\
	0 & : \text{otherwise}.
	\end{cases}
	\]

\medskip
\begin{enumerate}
\item Determine the value of $c$ that makes $f$ a valid density.\vfill
\item Compute $P(X\geq Y)$.\vfill
\end{enumerate}









\prob Suppose that the continuous random vector $(X,Y)$ is uniformly distributed over the triangle in $\mathbb{R}^2$ with vertices $(-1,0)$, $(1,0)$, and $(0,1)$.

\medskip
\begin{enumerate}
\item Compute the density function of $(X,Y)$.\vfill
\item Compute $P(X \leq 3/4, Y \leq 3/4)$.\vfill
\end{enumerate}

















\bigskip
\prob Suppose that $(X,Y)$ is continuous with probability density function

	\[
	f(x,y) = \begin{cases}
	30xy^2 & : x-1 \leq y \leq 1-x, \ 0 \leq x \leq 1, \\
	0 & : \text{otherwise}.
	\end{cases}
	\]

Compute $F(1/2,1/2)$.\vfill















\newpage
\prob Compute the marginal probability mass function distribution $p_X(x)$ of the discrete random vector $(X,Y)$ with probability mass function


	\[
	\begin{tabular}{c|c|c|c|c|c}
	$x \diagdown y$ & 0 & 1 & 2 & 3 & 4  \\ \hline
	0 & 0.08 & 0.07 & 0.06 & 0.01 & 0.01 \\ \hline
	1 & 0.06 & 0.10 & 0.12 & 0.05 & 0.02 \\ \hline
	2 & 0.05 & 0.06 & 0.09 & 0.04 & 0.03 \\ \hline
	3 & 0.02 & 0.03 & 0.03 & 0.03 & 0.04 
	\end{tabular}
	\]

\medskip
Verify that your computations are correct by making sure all marginal probabilities sum to $1$. How would you compute the other marginal mass function $p_Y(y)$?\vfill













\bigskip
\prob Suppose that $(X,Y)$ is continuous with probability density function

	\[
	f(x,y) = \begin{cases}
	\frac{21}{4}x^2 y & : x^2 \leq y \leq 1, \\
	0 & : \text{otherwise}.
	\end{cases}
	\]

Compute the marginal density functions $f_X(x)$ and $f_Y(y)$.\vfill






















\bigskip
\prob Suppose the joint PMF of two discrete random variables $X$ and $Y$ is given by

	\[
	\begin{tabular}{c|c|c|c|c}
	$x \diagdown y$ & 1 & 2 & 3 & 4 \\ \hline 
	1 & 0.1 & 0 & 0.1 & 0\\ \hline
	2 & 0.3 & 0 & 0.1 & 0.2 \\ \hline
	3 & 0 & 0.2 & 0 & 0
	\end{tabular}
	\]

Determine the conditional mass function $p_{X|Y}(x|1)$.\vfill
















\newpage
\prob Suppose that $(X,Y)$ is continuous with probability density function

	\[
	f(x,y) = \begin{cases}
	\frac{21}{4}x^2 y & : x^2 \leq y \leq 1, \\
	0 & : \text{otherwise}.
	\end{cases}
	\]

Compute the conditional density function $f_{Y|X}(y|x)$.\vfill














\bigskip
\prob A soft-drink machine has a random amount $Y$ in supply at the beginning of a given day and dispenses a random amount $X$ during the day (with measurements in gallons). It is not resupplied during the day, hence $X\leq Y$. It has been observed that $X$ and $Y$ have joint density given by

	\[
	f(x,y) = \begin{cases}
	1/2, & : 0 \leq x \leq y \leq 2, \\
	0 & : \text{otherwise}.
	\end{cases}
	\]

Find the conditional density $f(x|y)$ and evaluate the probability that less than $1/2$ gallon will be sold, given that the machine contains $1.5$ gallons at the start of the day.\vfill














\newpage
\prob Suppose that a person’s score $X$ on a mathematics aptitude test is a number between 0 and 1, and that their score $Y$ on a music aptitude test is also a number between 0 and 1. Suppose further that in the population of all college students in the United States, the scores $X$ and $Y$ are distributed according to the following joint PDF

	\[
	f(x,y) = \begin{cases}
	\frac{2}{5} (2x+3y) & : 0\leq x,y \leq 1, \\
	0 & : \text{otherwise}.
	\end{cases}
	\]

\medskip
\begin{enumerate}
\item What proportion of college students obtain a score greater than 0.8 on the mathematics test?\vfill
\item If a student’s score on the music test is 0.3, what is the probability that their score on the mathematics test will be greater than 0.8?\vfill
\item If a student’s score on the mathematics test is 0.3, what is the probability that their score on the music test will be greater than 0.8?\vfill
\end{enumerate}














\bigskip
\prob Let $X$ be the number of heads obtained from a single flip of a coin, so that $X\sim \mathcal{B}er(\theta)$ for some unknown probability $\theta$. Suppose further that $\theta$ is an observed value of a $\mathcal{B}eta(2,2)$ random variable. If we flip the coin and obtain $x=1$, how should we ``update'' the distribution of $\theta$?\vfill













\newpage
\prob Suppose that three random vectors $X$, $Y$, and $Z$ are jointly continuous with density function

	\[
	f(x,y,z) = \begin{cases}
	c(x+2y+3z) & : 0\leq x, y, z \leq 1, \\
	0 & : \text{otherwise}.
	\end{cases}
	\]

\medskip
\begin{enumerate}
\item Determine the value of $c$ that makes $f(x,y,z)$ a valid density function.\vfill
\item Compute the marginal density $f_{XY}(x,y)$.\vfill
\item Compute the probability $P\left( Z < 1/2 \mid X = 1/4,  Y = 3/4 \right)$.\vfill
\end{enumerate}









\bigskip
\prob Suppose that $X$, $Y$, and $Z$ have joint ``mixed density'' function

	\[
	f(x,y,z) = \begin{cases}
	c x^{1+y+z} ( 1-x)^{3-y-z} & : 0 < x< 1, \ y,z\in \{0,1\}, \\
	0 & : \text{otherwise}.
	\end{cases}
	\]
	
\medskip
\begin{enumerate}
\item Determine the value of $c$.\vfill
\item Compute the marginal ``density'' $f_{XY}(x,y)$.\vfill
\item Compute the conditional ``density'' $f_{Z|XY}(z|1/4, 1)$.\vfill
\end{enumerate}













\newpage
\prob Suppose that two measurements $X$ and $Y$ are made of the rainfall at a certain location on May 1 of two consecutive years. Supposing that $X$ and $Y$ are independent and that their marginal density functions are each given by

	\[
	f_X(x) = \begin{cases}
	2x & : 0 \leq x \leq 1, \\
	0 & : \text{otherwise},
	\end{cases} \quad \text{and} \quad
	f_Y(y) = \begin{cases}
	2y & : 0 \leq y \leq 1, \\
	0 & : \text{otherwise},
	\end{cases}
	\]

determine their joint density and compute the probability $P(X + Y \leq 1)$.\vfill


	
	
	
	
	
	















\bigskip
\prob Suppose that the joint density function of two continuous random variables $X$ and $Y$ is given by

	\[
	f(x,y) = \begin{cases}
	k x^2y^2 & : x^2+y^2 \leq 1, \\
	0 & : \text{otherwise},
	\end{cases}
	\]
	
for some constant $k$. Prove that $X$ and $Y$ are dependent.\vfill














\bigskip
\prob Suppose that a point $(X, Y )$ is chosen at random from the rectangle $R$ defined as follows:

	\[
	R=\{(x,y) : 0\leq x \leq 2, \ 1\leq y\leq 4\}.
	\]

\medskip
\begin{enumerate}
\item Determine the joint density of $X$ and $Y$, the marginal density of $X$, and the marginal density of $Y$.\vfill
\item Are $X$ and $Y$ independent?\vfill
\end{enumerate}






\newpage
\prob Let $n\geq 1$ be an integer and suppose $X \sim \mathcal{G}am(n+1,1)$. Suppose that $Y_1,Y_2,\ldots,Y_n$ is an identically distributed random sample that is independent given $X$ with density

	\[
	f(y|x) = \begin{cases}
	\frac{1}{x} & : 0 < y < x, \\
	0 & : \text{otherwise}.
	\end{cases}
	\]
	
\medskip
\begin{enumerate}
\item Determine the joint density of the random sample.\vfill
\item Determine the conditional density of $X$ for any given observed values of the random sample.\vfill
\end{enumerate}


\end{document}